---
author: V. Keith Hughitt
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
  settings: ""
  title: "Consensus Co-expression Network Construction"
  version: ""
  analysis_dirname: "03-consensus-network-construction"
title: "`r params$title` (`r params$version`)"
output:
  html_document:
    toc: true
    number_sections: true
  pdf_document:
    toc: true
    number_sections: true
    latex_engine: xelatex
---

Introduction
============

The goal of this analysis is to construct a consensus co-expression network by
summing the adjacency matrices for multiple network parameterizations, and then
compare the effect of pruning low-confidence network edges and genes at various
thresholds.

See the "Consensus co-expression network" analysis for a more thorough
explanation of the basic approach, and some alternative methods for
constructing the consensus network.

Methods
=======

```{r load_libraries, message = FALSE, results = 'hide'}
library(annotables)
library(biomaRt)
library(clusteval)
library(data.table)
library(digest)
library(doParallel)
library(plyr)
library(dynamicTreeCut)
library(DT)
library(flashClust)
library(foreach)
library(goseq)
library(gplots)
library(heatmaply)
library(hpgltools)
library(igraph)
library(knitr)
library(knitcitations)
library(reshape2)
library(tools)
library(tidyverse)

source('../../2015/00-shared/R/annotations.R')
source('../../2015/00-shared/R/enrichment_analysis.R')
source('../../2015/00-shared/R/filtering.R')
source('../../2015/00-shared/R/util.R')
source('../../2015/00-shared/R/wgcna.R')
```

```{r child = 'child/00-shared-setup.Rmd'}
```

```{r child = '../../2015/00-shared/Rmd/init/load_counts.Rmd'}
```

```{r child = '../../2015/00-shared/Rmd/init/load_host_annotations.Rmd', eval = CONFIG$target == 'host'}
```

```{r child = '../../2015/00-shared/Rmd/init/load_pathogen_annotations.Rmd', eval = CONFIG$target == 'pathogen'}
```

```{r child = 'child/03-build-consensus-net.Rmd'}
```

```{r child = 'child/03-consensus-net-functional-enrichment.Rmd'}
```

```{r child = 'child/03-filtered-consensus-net-comparison.Rmd'}
```

```{r child = 'child/03-filtered-consensus-net-results.Rmd'}
```

Results
=======

```{r child = 'child/03-consensus-net-vs-indiv-nets.Rmd'}
```

```{r child = 'child/03-consensus-net-vs-indiv-nets-human-grn.Rmd', eval = CONFIG$target == 'host'}
```

```{r save_tables}
# save tables to output dir
write.csv(table_1, file = file.path(output_tabledir, paste0('table_01_', output_suffix, '.csv')),
          quote = FALSE, row.names = FALSE)
```

System Information
==================

```{r sysinfo}
sessionInfo()
```

```{r git_commit}
# manuscript-specific code (https://github.com/elsayed-lab/manuscript-coex-networks)
system('git rev-parse --short HEAD', intern = TRUE)

# shared code (https://github.com/elsayed-lab/manuscript-shared-rnaseq)
system('git --git-dir=$UMD/2015/00-shared/.git rev-parse --short HEAD', intern = TRUE)
```

```{r save_output, include = FALSE}
# free up memory for pandoc conversion
rm(adjmat, filtered_adjmat)
suppressMessages(gc())

# Give rmarkdown some time to finish convern and then copy original rmarkdown
# along with output images and HTML to archive location
html_filepath <- file.path(output_prefix, paste0('network_construction_', MANUSCRIPT_CONFIG$output_prefix, '.html'))
pdf_filepath <- sub('html', 'pdf', html_filepath)

if (opts_knit$get("rmarkdown.pandoc.to") == 'latex') {
    system(sprintf('(sleep 60 && mv 03-consensus-network-construction.pdf %s) &', pdf_filepath))
} else {
    system(sprintf('(sleep 60 && mv 03-consensus-network-construction.html %s) &', html_filepath))
}
```

