## Overview

```{r results='asis'}
if (CONFIG$target == 'host' && CONFIG$host == 'H. sapiens') {
    # GO, KEGG, CPDB, TFs
    dat <- data.frame(
        cutoff_quantile=quantile_cutoffs,
        cutoff=cutoffs,
        num_modules,
        num_genes=genes_passing,
        median_module_size=median_module_sizes,
        largest_module_size=largest_module_sizes,
        num_unique_go_terms,
        num_unique_kegg_pathways,
        num_unique_cpdb_pathways,
        num_unique_tfs,
        total_go_pvals,
        total_kegg_pvals,
        total_cpdb_pvals,
        total_tf_pvals,
        num_go_enriched_modules,
        num_kegg_enriched_modules,
        num_tf_enriched_modules
    ) %>% mutate(
        `Ratio GO-enriched modules`=num_go_enriched_modules/num_modules,
        `Ratio KEGG-enriched modules`=num_kegg_enriched_modules/num_modules,
        `Ratio CPDB-enriched modules`=num_cpdb_enriched_modules/num_modules,
        `Ratio TF-enriched modules`=num_tf_enriched_modules/num_modules,
        `GO enrichment per gene`=total_go_pvals/num_genes,
        `KEGG enrichment per gene`=total_kegg_pvals/num_genes,
        `CPDB enrichment per gene`=total_cpdb_pvals/num_genes,
        `TF enrichment per gene`=total_tf_pvals/num_genes
    ) %>% dplyr::rename(
        `# Unique GO terms`=num_unique_go_terms,
        `# Unique KEGG pathways`=num_unique_kegg_pathways,
        `# Unique CPDB pathways`=num_unique_cpdb_pathways,
        `# Unique TF's`=num_unique_tfs,
        `GO enrichment -log10(p-val) score`=total_go_pvals,
        `KEGG enrichment -log10(p-val) score`=total_kegg_pvals,
        `CPDB enrichment -log10(p-val) score`=total_cpdb_pvals,
        `TF enrichment -log10(p-val) score`=total_tf_pvals
    )
} else {
    # GO, KEGG
    dat <- data.frame(
        cutoff_quantile=quantile_cutoffs,
        cutoff=cutoffs,
        num_modules,
        num_genes=genes_passing,
        median_module_size=median_module_sizes,
        largest_module_size=largest_module_sizes,
        num_unique_go_terms,
        num_unique_kegg_pathways,
        total_go_pvals,
        total_kegg_pvals,
        num_go_enriched_modules,
        num_kegg_enriched_modules
    ) %>% mutate(
        `Ratio GO-enriched modules`=num_go_enriched_modules/num_modules,
        `Ratio KEGG-enriched modules`=num_kegg_enriched_modules/num_modules,
        `GO enrichment per gene`=total_go_pvals/num_genes,
        `KEGG enrichment per gene`=total_kegg_pvals/num_genes,
    ) %>% dplyr::rename(
        `# Unique GO terms`=num_unique_go_terms,
        `# Unique KEGG pathways`=num_unique_kegg_pathways,
        `GO enrichment -log10(p-val) score`=total_go_pvals,
        `KEGG enrichment -log10(p-val) score`=total_kegg_pvals
    )
}

datatable(dat)
```

## Number of unique annotations

```{r ggplot_setup}
theme_set(theme_bw(base_size=16))
```

```{r num_unique_annotations, fig.width=1600/192, fig.height=1080/192}
dat_num_unique <- dat %>%
    select(cutoff_quantile, starts_with('# Unique')) %>%
    melt(id.vars='cutoff_quantile', variable.name='type',
         value.name='num_unique_annotations')

ggplot(dat_num_unique, aes(x=cutoff_quantile, y=num_unique_annotations, color=type)) +
    geom_line() +
    geom_vline(xintercept=MANUSCRIPT_CONFIG$quantile_cutoff, linetype='dashed', color='red') +
    theme_bw() +
    xlab('Edge Weight Cutoff Percentile') + ylab('# Unique annotations') +
    ggtitle('Number of Unique Annotations vs. Edge Weight Cutoff')
```

## -log10(Adj. P-value) totals

```{r pval_totals, fig.width=1600/192, fig.height=1080/192}
dat_pval_scores <- dat %>%
    select(cutoff_quantile, ends_with('score')) %>%
    melt(id.vars='cutoff_quantile', variable.name='type',
         value.name='pvalue_totals')

ggplot(dat_pval_scores, aes(x=cutoff_quantile, y=pvalue_totals, color=type)) +
    geom_line() +
    geom_vline(xintercept=MANUSCRIPT_CONFIG$quantile_cutoff, linetype='dashed', color='red') +
    theme_bw() +
    xlab('Edge Weight Cutoff Percentile') + ylab('Total -log10(Adj. P-values)') +
    ggtitle('Total -log10(Adj. P-values) vs. Edge Weight Cutoff')
```

## Ratio of enrichment score to number of genes

```{r ratio_enrichment_to_num_genes, fig.width=1600/192, fig.height=1080/192}
# sum(-log10(adj.p.val)) over # genes vs. cutoff
dat_ratio_enrichment <- dat %>%
    select(cutoff_quantile, ends_with('per gene')) %>%
    melt(id.vars='cutoff_quantile', variable.name='type',
         value.name='ratio_enrichment_to_num_genes')

ggplot(dat_ratio_enrichment, aes(x=cutoff_quantile, y=ratio_enrichment_to_num_genes, color=type)) +
    geom_line() +
    geom_vline(xintercept=MANUSCRIPT_CONFIG$quantile_cutoff, linetype='dashed', color='red') +
    theme_bw() +
    xlab('Edge Weight Cutoff Percentile') + ylab('Ratio Enrichment to Number of Genes') +
    ggtitle('Ratio of Enrichment to Number of Genes vs. Edge Weight Cutoff')
```

## Functional enrichment comparison (consensus vs. individual networks)

Next, we will compare the overall distributions of functional enrichment
between the robust consensus networks generated with various cutoffs, and
the unfiltered individual networks.

Note that because the individual networks are all unfiltered, they will have
more total genes (and thus more chances for enrichment) than many of the
consensus networks.

As such, we will only compare filtered consensus networks with half or more of
the original genes present.

```{r comparison_to_individual_networks, fig.width=1600/192, fig.height=1080/192}
nets <- MANUSCRIPT_CONFIG$networks

# compute score for each network using -log10 p-value scores, if available, or
# else total unique annotations

# score (-log10 p-value)
if ('total_go_pval' %in% colnames(nets)) {
    # total p-value scores by annotation type
    indiv_scores <- nets %>%
        select(starts_with('total'), id) %>%
        select(ends_with('pval'), id) %>%
        melt(id.vars='id', variable.name='type',
                value.name='total')

    if (CONFIG$target == 'host') {
        # human consensus network scores
        consensus_scores <- data.frame(
            type=c("GO Terms", "KEGG Pathways", "CPDB Pathways", "Marbach TF Genes"),
            score=c(dat$`GO enrichment -log10(p-val) score`[1], 
                    dat$`KEGG enrichment -log10(p-val) score`[1],
                    dat$`CPDB enrichment -log10(p-val) score`[1], 
                    dat$`TF enrichment -log10(p-val) score`[1])
        )

    } else {
        # parasite consensus network scores
        consensus_scores <- data.frame(
            type=c("GO Terms", "KEGG Pathways"),
            score=c(dat$`GO enrichment -log10(p-val) score`[1], 
                    dat$`KEGG enrichment -log10(p-val) score`[1])
        )
    }
    # replace labels
    label_mapping <- data.frame(
        from=c('total_go_pval', 'total_kegg_pval', 
                'total_cpdb_pval', 'total_marbach_pval', 'total_leishcyc_pval'),
        to=c("GO Terms", "KEGG Pathways", "CPDB Pathways", "Marbach TF Genes",
             "LeishCyc Pathways")
    )
    indiv_scores$type <- label_mapping$to[match(indiv_scores$type, label_mapping$from)]

} else {
    # total unique annotation counts by annotation type
    indiv_scores <- nets %>%
        select(starts_with('unique'), id) %>%
        melt(id.vars='id', variable.name='type',
                value.name='total')

    if (CONFIG$target == 'host') {
        # human consensus network scores
        consensus_scores <- data.frame(
            type=c("GO Terms", "KEGG Pathways", "CPDB Pathways", "Marbach TF Genes"),
            score=c(dat$`# Unique GO terms`[1], 
                    dat$`# Unique KEGG pathways`[1],
                    dat$`# Unique CPDB pathways`[1], 
                    dat$`# Unique TF's`[1])
        )

    } else {
        # parasite consensus network scores
        consensus_scores <- data.frame(
            type=c("GO Terms", "KEGG Pathways"),
            score=c(dat$`# Unique GO terms`[1], 
                    dat$`# Unique KEGG pathways`[1])
        )
    }
    # replace labels
    label_mapping <- data.frame(
        from=c('unique_go_terms', 'unique_kegg_terms', 
                'unique_cpdb_terms', 'unique_marbach_terms'),
        to=c("GO Terms", "KEGG Pathways", "CPDB Pathways", "Marbach TF Genes")
    )
    indiv_scores$type <- label_mapping$to[match(indiv_scores$type, label_mapping$from)]
}

# Exclude LeishCyc for now (sparesely annotated)
indiv_scores <- indiv_scores %>% 
    filter(type %in% c("GO Terms", "KEGG Pathways", "CPDB Pathways", "Marbach TF Genes"))

# plot of p-value scores by annotation type
ggplot(indiv_scores, aes(x=total, fill=type)) +
    geom_density() +
    geom_vline(data=consensus_scores, aes(xintercept=score), 
                linetype='dashed', color='red') +
    theme_bw() +
    facet_wrap(~ type, scale = 'free_x')

# Comparison of unfiltered consensus network score to distribution of
# individual network scores
#ggplot(data.frame(score=indiv_scores), aes(score)) +
#    geom_histogram()
    #geom_vline(xintercept=, linetype='dashed', color='red')

# Comparison of 0-50th percentile filtered consensus networks to all 
# individual networks
#if ('total_go_pval' %in% colnames(nets)) {
#    df <- rbind(data.frame(score=dat_pval_scores$pvalue_totals[1:51], type="Consensus Networks"),
#                data.frame(score=nets$total_go_pval, type='Individual Networks'))

#    ggplot(df, aes(x=score, group=type, fill=type)) +
#        geom_density(alpha=0.5)
#}

```

```{r table_1, results='asis'}
# comparison of unfiltered consensus network enrichment to distribution of
# individual networks
table_1 <- consensus_scores

quantiles <- c()

for (i in 1:nrow(table_1)) {
    type <- table_1$type[i]
    vals <- indiv_scores$total[indiv_scores$type == type]

    # quantile where consensus enrichment score falls in the individual
    # network distribution
    qval <- sum(table_1$score[i] > vals) / length(vals)
    quantiles <- c(quantiles, qval)
}

table_1 <- cbind(table_1, quantile=quantiles)

xkable(table_1, caption='Table 1. Comparison of unfiltered consensus network to individual networks')
```

