## Edge overlap

Next, we will randomly select N networks and measure the edge similarities
for each pair of networks.

We will evaluate edge similarity in two ways:

1. Measure the pearson correlation of adjacency matrices containing only shared genes
2. Measure the ratio of shared edges in the upper quantile of each network's edge weight
   distribution.

```{r compare_network_edge_weights, cache = CONFIG$use_cache, cache.lazy=FALSE, autodep=TRUE, message=FALSE}
# result vectors
output_adjmat_cor <- matrix(nrow=num_networks, ncol=num_networks)
output_shared_top_edges <- matrix(nrow=num_networks, ncol=num_networks)

rownames(output_adjmat_cor) <- colnames(output_adjmat_cor) <- paste0("network_", networks$id)
rownames(output_shared_top_edges) <- colnames(output_shared_top_edges) <- paste0("network_", networks$id)

# matrix to keep track of pairwise comparisons made 
nets_compared <- matrix(FALSE, nrow(networks), nrow(networks))

# iterate over pairs of networks
for (i in 1:nrow(networks)) {
  # load first network
  id_a = networks$id[i]
  adjmat_file_a <- sprintf("%d_adjmat.RData", id_a)
  mapping_file_a <- sprintf("%d.RData", id_a)

  load(file.path(MANUSCRIPT_CONFIG$input_dir, mapping_file_a))
  load(file.path(MANUSCRIPT_CONFIG$input_dir, adjmat_file_a))

  # edge count matrix (upper triangular)
  module_mapping_a <- module_mapping
  matsize_a <- nrow(module_mapping_a)
  genes_a <- module_mapping_a$gene_id

  adjmat_a_unfiltered <- matrix(0, nrow=matsize_a, ncol=matsize_a)
  adjmat_a_unfiltered[upper.tri(adjmat_a_unfiltered)] <- adjacency_matrix

  for (j in 1:nrow(networks)) {
    # skip self-comparisons and symmetric comparisons
    if (i == j || nets_compared[i, j]) {
      next
    }

    message(sprintf("Comparing networks %d vs. %d...", i, j))

    # load second network
    id_b = networks$id[j]
    adjmat_file_b <- sprintf("%d_adjmat.RData", id_b)
    mapping_file_b <- sprintf("%d.RData", id_b)

    load(file.path(MANUSCRIPT_CONFIG$input_dir, mapping_file_b))
    load(file.path(MANUSCRIPT_CONFIG$input_dir, adjmat_file_b))

    # adjacency matrix (upper triangular)
    module_mapping_b <- module_mapping
    matsize_b <- nrow(module_mapping_b)
    genes_b <- module_mapping_b$gene_id

    adjmat_b <- matrix(0, nrow=matsize_b, ncol=matsize_b)
    adjmat_b[upper.tri(adjmat_b)] <- adjacency_matrix

    # drop any genes only present in one network (differences may exist due to differing batch
    # adjustment settings, even if all filtering criteria are held constant)
    mask_a <- genes_a %in% genes_b
    mask_b <- genes_b %in% genes_a

    # TRUE
    # all(module_mapping_a$gene_id[mask_a] == module_mapping_b$gene_id[mask_b])

    # create a copy of adjmat_a and update (we want to leave original adjmat_a untouched since it
    # will be compared against other networks in subsequent loop iterations..)
    adjmat_a <- adjmat_a_unfiltered

    adjmat_a <- adjmat_a[mask_a, mask_a]
    adjmat_b <- adjmat_b[mask_b, mask_b]

    # measure correlation of matrix upper triangles
    adj_cor <- cor(adjmat_a[upper.tri(adjmat_a)], adjmat_b[upper.tri(adjmat_b)])
    output_adjmat_cor[i, j] <- output_adjmat_cor[j, i] <- adj_cor

    # next, binarize each matrix so that only the edges in the top 90th percentile of each
    # network's edge weight distribution remain and are set to 1
    edge_cutoff_a <- quantile(adjmat_a[upper.tri(adjmat_a)], EDGE_WEIGHT_QUANTILE_CUTOFF)
    edge_cutoff_b <- quantile(adjmat_b[upper.tri(adjmat_b)], EDGE_WEIGHT_QUANTILE_CUTOFF)

    # measure overlap of binarized versions of each matrix
    binary_mat_b <- adjmat_b[upper.tri(adjmat_b)] > edge_cutoff_b
    rm(adjmat_b)
    suppressMessages(gc())

    binary_mat_a <- adjmat_a[upper.tri(adjmat_a)] > edge_cutoff_a

    # ratio of shared top edges
    shared_ratio <- sum(binary_mat_a & binary_mat_b) / sum(binary_mat_a | binary_mat_b)
    output_shared_top_edges[i, j] <- output_shared_top_edges[j, i] <- shared_ratio

    # update tracking matrix
    nets_compared[i, j] <- TRUE
    nets_compared[j, i] <- TRUE
  }
} 

# Set output diagonals to maximum of off-diagonal values to increase dynamic range of color scale
diag(output_adjmat_cor) = 0
diag(output_adjmat_cor) = max(output_adjmat_cor)

diag(output_shared_top_edges) = 0
diag(output_shared_top_edges) = max(output_shared_top_edges)

# clean-up
rm(adjmat_a_unfiltered)
rm(adjmat_a)
suppressMessages(gc())
```

### Network edge weight correlations

#### Distribution of edge weight correlations

```{r network_edgeweight_correlation_dist}
plot(density(output_adjmat_cor[upper.tri(output_adjmat_cor)]), 
     main='Distribution of edge weight correlations', xlab='Correlation')
```

#### Heatmap of edge weight correlations

```{r network_edgeweight_correlation_heatmap}
heatmap.2(output_adjmat_cor, trace='none', dendrogram='row', margin=c(4, 4),
          cexRow=heatmap_cex, cexCol=heatmap_cex, col='plasma',
          key.title=NA, key.xlab='Pearson Correlation',
          symbreaks=FALSE, symkey=FALSE)
```

### Ratio of shared edges

#### Distribution of ratios of shared top edge-weights

```{r network_ratio_shared_top_edges_dist}
plot(density(output_shared_top_edges[upper.tri(output_shared_top_edges)]), 
     main='Distribution of ratios of shared top edge-weights', xlab='Ratio')
```

```{r network_ratio_shared_top_edges_heatmap}
heatmap.2(output_shared_top_edges, trace='none', dendrogram='row', margin=c(4, 4),
          cexRow=heatmap_cex, cexCol=heatmap_cex, col='plasma',
          key.title=NA, key.xlab='Ratio Shared Top Edges',
          symbreaks=FALSE, symkey=FALSE)
```
