## Consensus co-expression network construction

To begin with, previously generated adjaceny matrices from multiple different
network parameterizations are loaded and combined into a single "consensus
adjacency matrix".

Each cell in the matrix represents the total support for a co-expression
relationship between two genes across a large number of alternative network
instantiations.

```{r load_consensus_nets, message=FALSE, cache = CONFIG$use_cache, cache.lazy=FALSE, autodep=TRUE}
# input adjaceny matrices
input_nets <- Sys.glob(file.path(MANUSCRIPT_CONFIG$input_dir, '*_adjmat.RData'))
input_rdas <- sub('_adjmat', '', input_nets)

if (length(input_nets) == 0) {
  stop("No input networks found! check to make sure paths are correctly defined.")
}

# filter input nets and rdas (optional)
if ('networks' %in% names(MANUSCRIPT_CONFIG)) {
    nets_to_include <- sprintf('%d_adjmat.RData', MANUSCRIPT_CONFIG$networks$id)
    rdas_to_include <- sprintf('%d.RData', MANUSCRIPT_CONFIG$networks$id)

    input_nets <- input_nets[basename(input_nets) %in% nets_to_include] 
    input_rdas <- input_rdas[basename(input_rdas) %in% rdas_to_include] 
}

# empty vectors to keep track of network metrics
ratio_nonzero_edges  <- c()
ratio_edges_1q       <- c()
ratio_edges_2q       <- c()
ratio_edges_3q       <- c()
median_edge_weight_global   <- c()

# Note: for performance purposes, we will hold off converting the stored upper triangular 
# adjacency matrix values until the very last step. 
# We will limit ourselves to only those genes found in all networks; currently,
# there are few parameters which affect the total number of genes in the network, and as such,
# there are likely only a few possible combinations of genes present. To avoid having to convert
# the adj mat upper triangular values back into a matrix (slow for large adjacency matrices) each 
# time to determine which entries need to be removed, we will instead maintain multiple separate 
# "sub" consensus vectors; one for each gene subset present. After all of the matrices have been
# loaded, we will then convert each configuration back in to a matrix and remove genes not
# present in all networks, before combining into a final consensus network.

# Example (HsTc v7.0)
# > table(x$num_genes)
# 29109 29148 29219 
#  224   224   896 
#

# lists to keep track of sub-consensus networks associated with each gene configuration present
# (typically only one to three such configurations present for a given dataset)
adjmats <- list()
adjmat_genes <- list()

# get individual network scores
networks <- MANUSCRIPT_CONFIG$networks

if (CONFIG$target == 'host') {
    # Host (GO, CPDB, Marbach)
    #networks$enrichment_score = as.vector((1/3) * (scale(networks$total_go_pval) +
    #                                               scale(networks$total_cpdb_pval) +
    #                                               scale(networks$total_marbach_pval)))
		networks$enrichment_score = as.vector(networks$total_go_pval + networks$total_cpdb_pval + networks$total_marbach_pval)
} else {
    # Parasite (GO)
    networks$enrichment_score = as.vector(scale(networks$total_go_pval))
}

net_ids <- as.numeric(str_match(input_nets, '([0-9]+)_adjmat.RData')[, 2])
net_scores <- networks$enrichment_score[match(net_ids, networks$id)]

# iterate over remaining networks and add to consensus net
#
for (i in 1:length(input_nets)) {
    # load saved data
    message(sprintf("Loading network %d/%d (%s)", i, length(input_nets), basename(input_nets[i])))
    message(timestamp())

    load(input_nets[i])
    load(input_rdas[i])

    # check to make sure dimensions are as expected (a few adjacency matrices
    # in the past were corrupted, leading to issues; the problem has since been fixed
    # upstream and should no longer occur, but leaving check just to be safe..)
    matsize <- nrow(module_mapping)

    if ((((matsize * matsize) - matsize) / 2) != length(adjacency_matrix)) {
        warning(sprintf("Dimension mismatch for adjacency matrix: %s", input_rdas[i]))
        stop(sprintf("Incompatible adjacency matrix input: %s", input_nets[i]))
    }

    # ratio of non-zero edges (global)
    ratio_nonzero_edges <- c(ratio_nonzero_edges, sum(adjacency_matrix != 0) /
                             length(adjacency_matrix))

    # ratio of edges with values > 0.25, 0.50, 0.75 of the maximum edge weight
    scaled_adjmat <- adjacency_matrix / max(adjacency_matrix)

    ratio_edges_1q <- c(ratio_edges_1q, sum(scaled_adjmat > 0.25) / length(scaled_adjmat))
    ratio_edges_2q <- c(ratio_edges_2q, sum(scaled_adjmat > 0.50) / length(scaled_adjmat))
    ratio_edges_3q <- c(ratio_edges_3q, sum(scaled_adjmat > 0.75) / length(scaled_adjmat))

    median_edge_weight_global <- c(median_edge_weight_global, 
                                   median(adjacency_matrix[adjacency_matrix != 0]))

    #
    # normalize edge weights
    #
    # none  - do nothing
    # scale - size-factor scaling; adjmats all sum to 1E6
    # rank  - rank-based scaling
    # score - score-weighted scaling; size-factor scaling is applied and each individual network's
    #         contribution is a function of its score.
    #
    if ('normalization' %in% names(MANUSCRIPT_CONFIG)) {
        if (MANUSCRIPT_CONFIG$normalization == 'scale') {
            # size-factor normalization
            adjacency_matrix <- (adjacency_matrix / sum(adjacency_matrix)) * 1E6
        } else if (MANUSCRIPT_CONFIG$normalization == 'rank') {
            # rank-based normalization
            adjacency_matrix <- data.table::frank(adjacency_matrix) * 1E-5
        } else if (MANUSCRIPT_CONFIG$normalization == 'score') {
            # first, perform size-factor normalization
            adjacency_matrix <- (adjacency_matrix / sum(adjacency_matrix)) * 1E6

						# next, weight the contribution of the adjacency matrix based on its score and
						# divide by a scaling factor for convenience
						adjacency_matrix <- (adjacency_matrix * net_scores[i]) / 10
        }
    }

    # add to consensus sub-matrix (using md5 hash of gene ids to identify subset)
    key <- digest(module_mapping$gene_id, algo = 'md5')

    if (!key %in% names(adjmats)) {
        # new gene configuration
        adjmats[[key]]      <- adjacency_matrix
        adjmat_genes[[key]] <- module_mapping$gene_id
    } else {
        # existing gene configuration
        adjmats[[key]] <- adjmats[[key]] + adjacency_matrix
    }
}
```

```{r recombine_adj_submats}
# determine which genes are present in all networks
shared_genes <- Reduce(intersect, adjmat_genes)

# create empty adjacency matrix with expected dimensions
adjmat <- matrix(0, nrow = length(shared_genes), ncol = length(shared_genes))

rownames(adjmat) <- shared_genes
colnames(adjmat) <- shared_genes

# iterate over gene configurations and add to combined (final) consensus adjacency matrix
for (key in names(adjmats)) {
    # get gene ids (adjmats and module_mappings are always stored with rows/cols sorted by gene id)
    submat_genes <- adjmat_genes[[key]]

    # convert upper triangular matrix values back into full matrix
    matsize <- length(submat_genes) 

    # convert upper.tri weight vector back into adj matrix (normalized)
    tmp <- matrix(0, nrow = matsize, ncol = matsize)
    tmp[upper.tri(tmp)] <- adjmats[[key]]

    rownames(tmp) <- submat_genes
    colnames(tmp) <- submat_genes

    # drop non-shared genes, if present
    mask <- rownames(tmp) %in% shared_genes
    tmp <- tmp[mask, mask]

    # rows and columsn should already be in sorted order, but just to be safe...
    if (!all(rownames(tmp) == rownames(adjmat))) {
        stop("Mismatched adjacency matrix entries!")
    }

    # add to combined matrix
    adjmat <- adjmat + tmp
}

# fill in lower triangular part of adjacency matrix (symmetric)
adjmat[lower.tri(adjmat)] <- t(adjmat)[lower.tri(adjmat)]

# histogram of the consensus adj matrix edge weight distribution
#hist(log1p(adjmat))
sampled_adjmat <- sample(adjmat[lower.tri(adjmat)], 1E5)

hist(sampled_adjmat, main="Distribution of consensus adjacency matrix edge weights", xlab="Edge weight")
hist(sampled_adjmat, main="Distribution of consensus adjacency matrix edge weights (Log)", xlab="Log(Edge weight)")
```

```{r save_rdata}
# save RData files to output dir
save(adjmat, file=file.path(output_datadir, paste0('adjmat_', output_suffix, '.rda')))
rm(adjmats, sampled_adjmat, tmp)
suppressMessages(gc())
```

