## Functional enrichment comparison (consensus vs. individual networks)

Next, we will compare the overall distributions of functional enrichment
between the robust consensus networks generated with various cutoffs, and
the unfiltered individual networks.

Note that because the individual networks are all unfiltered, they will have
more total genes (and thus more chances for enrichment) than many of the
consensus networks.

As such, we will only compare filtered consensus networks with half or more of
the original genes present.

```{r comparison_to_individual_networks, fig.height=ifelse(CONFIG$target == 'host', 3.375, 2.5)}
nets <- MANUSCRIPT_CONFIG$networks

# compute score for each network using -log10 p-value scores, if available, or
# else total unique annotations

# score (-log10 p-value)
# total p-value scores by annotation type
indiv_scores <- nets %>%
    select(starts_with('total'), id) %>%
    select(ends_with('pval'), id) %>%
    melt(id.vars='id', variable.name='type',
            value.name='total')

if (CONFIG$target == 'host') {
    # human consensus network scores
    consensus_scores <- data.frame(
        type=c("GO Terms", "KEGG Pathways", "CPDB Pathways", "Marbach TF Genes"),
        score=c(go_pval_total, kegg_pval_total, cpdb_pval_total, tf_pval_total)
    )

} else {
    # parasite consensus network scores
    consensus_scores <- data.frame(
        type=c("GO Terms", "KEGG Pathways"),
        score=c(go_pval_total, 
                kegg_pval_total)
    )
}
# replace labels
label_mapping <- data.frame(
    from=c('total_go_pval', 'total_kegg_pval', 
            'total_cpdb_pval', 'total_marbach_pval', 'total_leishcyc_pval'),
    to=c("GO Terms", "KEGG Pathways", "CPDB Pathways", "Marbach TF Genes",
            "LeishCyc Pathways")
)
indiv_scores$type <- label_mapping$to[match(indiv_scores$type, label_mapping$from)]

# total unique annotation counts by annotation type
#indiv_scores <- nets %>%
#    select(starts_with('unique'), id) %>%
#    melt(id.vars='id', variable.name='type',
#            value.name='total')

#if (CONFIG$target == 'host') {
#    # human consensus network scores
#    consensus_scores <- data.frame(
#        type=c("GO Terms", "KEGG Pathways", "CPDB Pathways", "Marbach TF Genes"),
#        score=c(dat$`# Unique GO terms`[1], 
#                dat$`# Unique KEGG pathways`[1],
#                dat$`# Unique CPDB pathways`[1], 
#                dat$`# Unique TF's`[1])
#    )

#} else {
#    # parasite consensus network scores
#    consensus_scores <- data.frame(
#        type=c("GO Terms", "KEGG Pathways"),
#        score=c(dat$`# Unique GO terms`[1], 
#                dat$`# Unique KEGG pathways`[1])
#    )
#}
## replace labels
#label_mapping <- data.frame(
#    from=c('unique_go_terms', 'unique_kegg_terms', 
#            'unique_cpdb_terms', 'unique_marbach_terms'),
#    to=c("GO Terms", "KEGG Pathways", "CPDB Pathways", "Marbach TF Genes")
#)
#indiv_scores$type <- label_mapping$to[match(indiv_scores$type, label_mapping$from)]
##}

## Exclude LeishCyc for now (sparesely annotated)
#indiv_scores <- indiv_scores %>% 
#    filter(type %in% c("GO Terms", "KEGG Pathways", "CPDB Pathways", "Marbach TF Genes"))

# plot of p-value scores by annotation type
ggplot(indiv_scores, aes(x=total, fill=type)) +
    geom_density() +
    geom_vline(data=consensus_scores, aes(xintercept=score), 
                linetype='dashed', color='red') +
    theme_bw_high_res() +
    facet_wrap(~ type, scale = 'free_x')

# Comparison of unfiltered consensus network score to distribution of
# individual network scores
#ggplot(data.frame(score=indiv_scores), aes(score)) +
#    geom_histogram()
    #geom_vline(xintercept=, linetype='dashed', color='red')

# Comparison of 0-50th percentile filtered consensus networks to all 
# individual networks
#if ('total_go_pval' %in% colnames(nets)) {
#    df <- rbind(data.frame(score=dat_pval_scores$pvalue_totals[1:51], type="Consensus Networks"),
#                data.frame(score=nets$total_go_pval, type='Individual Networks'))

#    ggplot(df, aes(x=score, group=type, fill=type)) +
#        geom_density(alpha=0.5)
#}

```

```{r table_1, results='asis'}
# comparison of unfiltered consensus network enrichment to distribution of
# individual networks
table_1 <- consensus_scores

quantiles <- c()

for (i in 1:nrow(table_1)) {
    type <- table_1$type[i]
    vals <- indiv_scores$total[indiv_scores$type == type]

    # quantile where consensus enrichment score falls in the individual
    # network distribution
    qval <- sum(table_1$score[i] > vals) / length(vals)
    quantiles <- c(quantiles, qval)
}

table_1 <- cbind(table_1, quantile=quantiles)

xkable(table_1, caption='Table 1. Comparison of unfiltered consensus network to individual networks')
```

## Network metrics (individual vs. consensus networks)

```{r}
# create a max-scaled version of consensus network for a more equal comparison
# to individual nets
scaled_adjmat <- adjmat / max(adjmat)
```

### Network densities

```{r network_density}
# consensus network density
g <- graph.adjacency(scaled_adjmat, mode='undirected', weighted=TRUE)
consensus_val <- edge_density(g)
rm(g)

ggplot(data.frame(x=network_densities), aes(x)) +
    geom_density(fill='#7ebd37') +
    geom_vline(xintercept=consensus_val, linetype='dashed', color='#bd373b') +
    xlab("Network Density") +
    theme_bw_high_res()
```

## Ratio of non-zero edges (global)

```{r ratio_nonzero_edges_global}
consensus_val <- sum(scaled_adjmat != 0) / length(scaled_adjmat)

ggplot(data.frame(x=ratio_nonzero_edges), aes(x)) +
    geom_density(fill='#7ebd37') +
    geom_vline(xintercept=consensus_val, linetype='dashed', color='#bd373b') +
    xlab("Ratio of Non-zero Edges (global)") +
    theme_bw_high_res()
```

## Ratio of edges with values > 0.25

```{r ratio_edges_gt_1q}
consensus_val <- sum(scaled_adjmat > 0.25) / length(scaled_adjmat)

ggplot(data.frame(x=ratio_edges_1q), aes(x)) +
    geom_density(fill='#7ebd37') +
    geom_vline(xintercept=consensus_val, linetype='dashed', color='#bd373b') +
    xlab("Ratio of Edges > 0.25 (global)") +
    theme_bw_high_res()
```

## Ratio of edges with values > 0.50

```{r ratio_edges_gt_2q}
consensus_val <- sum(scaled_adjmat > 0.5) / length(scaled_adjmat)

ggplot(data.frame(x=ratio_edges_2q), aes(x)) +
    geom_density(fill='#7ebd37') +
    geom_vline(xintercept=consensus_val, linetype='dashed', color='#bd373b') +
    xlab("Ratio of Edges > 0.50 (global)") +
    theme_bw_high_res()
```

## Ratio of edges with values > 0.75

```{r ratio_edges_gt_3q}
consensus_val <- sum(scaled_adjmat > 0.75) / length(scaled_adjmat)

ggplot(data.frame(x=ratio_edges_3q), aes(x)) +
    geom_density(fill='#7ebd37') +
    geom_vline(xintercept=consensus_val, linetype='dashed', color='#bd373b') +
    xlab("Ratio of Edges > 0.75 (global)") +
    theme_bw_high_res()
```

## Median non-zero edge weight (global)

```{r median_nonzero_edge_weight_global}
consensus_val <- median(scaled_adjmat[scaled_adjmat != 0])

ggplot(data.frame(x=median_edge_weight_global), aes(x)) +
    geom_density(fill='#7ebd37') +
    geom_vline(xintercept=consensus_val, linetype='dashed', color='#bd373b') +
    xlab("Median Edge Weight (global)") +
    theme_bw_high_res()
```

## Median ratio of non-zero edges per gene

```{r median_ratio_nonzero_edges_per_gene}
consensus_val <- median(apply(scaled_adjmat, 1, function(x) { sum(x != 0) / nrow(scaled_adjmat) }))

ggplot(data.frame(x=median_ratio_edges_per_gene), aes(x)) +
    geom_density(fill='#7ebd37') +
    geom_vline(xintercept=consensus_val, linetype='dashed', color='#bd373b') +
    xlab("Median Ratio Non-zero Edges Per Gene") +
    theme_bw_high_res()
```

## Median Non-zero Edge Weight per Gene

```{r median_nonzero_edge_weight_per_gene}
consensus_val <- median(apply(scaled_adjmat, 1, function(x) { median(x[x != 0]) }))

ggplot(data.frame(x=median_edge_weight_per_gene), aes(x)) +
    geom_density(fill='#7ebd37') +
    geom_vline(xintercept=consensus_val, linetype='dashed', color='#bd373b') +
    xlab("Median Non-zero Edge Weight Per Gene") +
    theme_bw_high_res()
```
