## Functional enrichment comparison (consensus vs. individual networks)

Next, we will compare level of functional enrichment in the _unfiltered_ consensus network to the
distribution of enrichment scores for the various _unfiltered_ individual co-expression networks.

While it may be useful to use a filtered version of the consensus network for downstream analyses,
in order to allow for a fair comparison with the individual networks, only unfiltered networks
are shown below.

For each figure, the distribution of _individual_ network enrichment scores are shown. The red
dashed line indicates the level of enrichment for the _consensus_ network.

```{r comparison_to_individual_networks, fig.height=ifelse(CONFIG$target == 'host', 3.375, 2.5)}
nets <- MANUSCRIPT_CONFIG$networks

# if filtering is disabled, create necessary single-element vectors by hand
if (!'total_go_pvals' %in% ls()) {
    total_go_pvals   <- go_pval_total
    total_kegg_pvals <- kegg_pval_total

    if (CONFIG$target == 'host' && CONFIG$host == 'H. sapiens') {
        total_cpdb_pvals <- cpdb_pval_total
        total_tf_pvals   <- tf_pval_total
    }
}

# compute score for each network using -log10 p-value scores, if available, or
# else total unique annotations

# score (-log10 p-value)
# total p-value scores by annotation type
indiv_scores <- nets %>%
    select(starts_with('total'), id) %>%
    select(ends_with('pval'), id) %>%
    melt(id.vars='id', variable.name='type',
            value.name='total')

if (CONFIG$target == 'host') {
    # human consensus network scores
    consensus_scores <- data.frame(
        type=c("GO Terms", "KEGG Pathways", "CPDB Pathways", "Marbach TF Genes"),
        #score=c(go_pval_total, kegg_pval_total, cpdb_pval_total, tf_pval_total)
        score=c(total_go_pvals[1], total_kegg_pvals[1], total_cpdb_pvals[1], total_tf_pvals[1])
    )

} else {
    # parasite consensus network scores
    consensus_scores <- data.frame(
        type=c("GO Terms", "KEGG Pathways"),
        #score=c(go_pval_total, kegg_pval_total)
        score=c(total_go_pvals[1], total_kegg_pvals[1])
    )
}
# replace labels
label_mapping <- data.frame(
    from=c('total_go_pval', 'total_kegg_pval', 'total_cpdb_pval', 'total_marbach_pval'),
    to=c("GO Terms", "KEGG Pathways", "CPDB Pathways", "Marbach TF Genes")
)
indiv_scores$type <- label_mapping$to[match(indiv_scores$type, label_mapping$from)]

# plot of p-value scores by annotation type
ggplot(indiv_scores, aes(x=total, fill=type)) +
    geom_density() +
    geom_vline(data=consensus_scores, aes(xintercept=score), 
                linetype='dashed', color='red') +
    theme_bw_high_res() +
    facet_wrap(~ type, scale = 'free_x')
```

```{r table_1, results='asis'}
# comparison of unfiltered consensus network enrichment to distribution of
# individual networks
table_1 <- consensus_scores

quantiles <- c()

for (i in 1:nrow(table_1)) {
    type <- table_1$type[i]
    vals <- indiv_scores$total[indiv_scores$type == type]

    # quantile where consensus enrichment score falls in the individual
    # network distribution
    qval <- sum(table_1$score[i] > vals) / length(vals)
    quantiles <- c(quantiles, qval)
}

table_1 <- cbind(table_1, quantile=quantiles)

xkable(table_1, caption='Table 1. Comparison of unfiltered consensus network to individual networks')
```

## Network metrics (individual vs. consensus networks)

```{r}
# create a max-scaled version of consensus network for comparison to individual nets
scaled_adjmat <- adjmat / max(adjmat)
```

## Ratio of non-zero edges (global)

```{r ratio_nonzero_edges_global}
consensus_val <- sum(scaled_adjmat != 0) / length(scaled_adjmat)

ggplot(data.frame(x=ratio_nonzero_edges), aes(x)) +
    geom_density(fill='#7ebd37') +
    geom_vline(xintercept=consensus_val, linetype='dashed', color='#bd373b') +
    xlab("Ratio of Non-zero Edges (global)") +
    theme_bw_high_res()
```

## Ratio of edges with values > 0.25

```{r ratio_edges_gt_1q}
consensus_val <- sum(scaled_adjmat > 0.25) / length(scaled_adjmat)

ggplot(data.frame(x=ratio_edges_1q), aes(x)) +
    geom_density(fill='#7ebd37') +
    geom_vline(xintercept=consensus_val, linetype='dashed', color='#bd373b') +
    xlab("Ratio of Edges > 0.25 (global)") +
    theme_bw_high_res()
```

## Ratio of edges with values > 0.50

```{r ratio_edges_gt_2q}
consensus_val <- sum(scaled_adjmat > 0.5) / length(scaled_adjmat)

ggplot(data.frame(x=ratio_edges_2q), aes(x)) +
    geom_density(fill='#7ebd37') +
    geom_vline(xintercept=consensus_val, linetype='dashed', color='#bd373b') +
    xlab("Ratio of Edges > 0.50 (global)") +
    theme_bw_high_res()
```

## Ratio of edges with values > 0.75

```{r ratio_edges_gt_3q}
consensus_val <- sum(scaled_adjmat > 0.75) / length(scaled_adjmat)

ggplot(data.frame(x=ratio_edges_3q), aes(x)) +
    geom_density(fill='#7ebd37') +
    geom_vline(xintercept=consensus_val, linetype='dashed', color='#bd373b') +
    xlab("Ratio of Edges > 0.75 (global)") +
    theme_bw_high_res()
```

## Median non-zero edge weight (global)

```{r median_nonzero_edge_weight_global}
consensus_val <- median(scaled_adjmat[scaled_adjmat != 0])

ggplot(data.frame(x=median_edge_weight_global), aes(x)) +
    geom_density(fill='#7ebd37') +
    geom_vline(xintercept=consensus_val, linetype='dashed', color='#bd373b') +
    xlab("Median Edge Weight (global)") +
    theme_bw_high_res()
```

```{r cleanup, include = FALSE, echo = FALSE}
rm(scaled_adjmat)
suppressMessages(gc())
```

