### Conserved infection signature network

Next, we will construct a "conservation matrix" by scoring each edge based on
its level of conservation across the networks infected networks. Edges with
large values in both infected networks and low values in both uninfected
networks receieve a high conservation score.

After rescaling each input network to the range [0, 1], the score is defined as:

$$
\text{score} = \frac{1}{2}(\text{HsLm + HsTc}) - \frac{1}{2}(\text{HsLmUI + HsTcUI})
$$

```{r generate_conservation_score_mat}
i1 <- hslm_inf / max(hslm_inf)
i2 <- hstc_inf / max(hstc_inf)
u1 <- hslm_uninf / max(hslm_uninf)
u2 <- hstc_uninf / max(hstc_uninf)

#score_mat <- pmin(i1, i2) - pmax(u1, u2)
score_mat <- 0.5 * (i1 + i2) - 0.5 * (u1 + u2)

hist(score_mat, main="Distribution of co-expression conservation scores")
```

```{r infection_sig_heatmap}
# sub-sample to speed up and better fit in display
ind <- sample(1:nrow(score_mat), 250)

heatmap.2(score_mat[ind, ind], trace='none', dendrogram='row', labRow=FALSE,
          labCol=FALSE, col='viridis',
          margin=c(6, 6), xlab='Gene', ylab='Gene', density='none',
          #key=FALSE,
          key.title='', key.ylab='',  keysize=2,
          key.xlab='Co-ex Conservation Score\n (HsTc / HsLm)')

# edge weight range
cat(sprintf('\n- Minimum edge weight: %0.3f\n', min(score_mat)))
cat(sprintf('\n- Maximum edge weight: %0.3f\n', max(score_mat)))
```

```{r score_matrix_clustering}
# since we are primarily interested in genes which become more strongly
# co-regulated during infection at the moment, we will clip at 0.
#score_mat[score_mat < 0] <- 0

# convert to distance matrix and perform hierarchical clustering
# use abs(score_mat)?
distmat <- 1 - ((1 + score_mat) / 2)

gene_tree <- flashClust(as.dist(distmat), method="average")
gene_tree$height <- round(gene_tree$height, 6)

# perform dynamic tree-cut based clustering
module_labels <- cutreeDynamicTree(dendro=gene_tree, deepSplit=FALSE,
                                   minModuleSize=10)

num_labels <- length(unique(module_labels))
pal <- c('grey', substring(rainbow(num_labels), 1, 7))
module_colors <- (pal[module_labels + 1])

module_mapping <- tbl_df(cbind(gene_id=rownames(score_mat), module=module_colors))

# create a mapping from module id to number of genes for later use
module_sizes <- setNames(as.data.frame(module_colors), 'module_id') %>% 
    group_by(module_id) %>% 
    summarize(num_genes=n())

num_modules <- length(unique(module_colors))
module_order <- unique(module_colors[gene_tree$order])
```

```{r module_stats, results='asis'}
cat(sprintf("- Number of modules: %d\n", num_modules))
cat(sprintf("- Median module size: %0.1f\n", median(module_sizes$num_genes)))
cat(sprintf("- Largest module size: %d\n", max(module_sizes$num_genes)))
```

### Functional Enrichment

```{r create_net_annotation_df}
# create dataframe to use for network node annotations
gene_info <- gene_info_orig %>% 
    filter(gene_id %in% rownames(score_mat))

result <- cbind(gene_info, color=module_colors)
annot <- cbind(gene_info, color=module_colors, cluster=module_labels)
```

#### GO Enrichment

```{r child='../../../2015/00-shared/Rmd/results/go_enrichment_network.Rmd'}
```

#### KEGG Enrichment

```{r child='../../../2015/00-shared/Rmd/results/kegg_enrichment_network.Rmd'}
```

```{r child='04-kegg-enrichment-results.Rmd'}
```

#### CPDB Enrichment

```{r child='../../../2015/00-shared/Rmd/results/cpdb_enrichment_network.Rmd'}
```

```{r child='04-kegg-enrichment-results.Rmd'}
```

#### TF regulon Enrichment

```{r child='../../../2015/00-shared/Rmd/results/hsapiens_marbach2016_tf_enrichment.Rmd'}
```

```{r child='04-tf-regulon-enrichment-results.Rmd'}
```

#### Visualization of enriched conserved subnetworks

```{r enriched_conservation_subnetwork, results='asis'}
cat('\n')

# heatmap color palette
pal <- colorRampPalette(c('slateblue1', 'coral1'))

for (module_color in unique(c(enriched_colors_go, enriched_colors_kegg,
                              enriched_colors_cpdb, enriched_colors_marbach))) {
    # show conservation network
    genes <- (result %>% filter(color==module_color))$gene_id
    ind <- rownames(score_mat) %in% genes

    cat(sprintf("###### %s (median score = %0.2f)\n", module_color, median(score_mat[ind, ind])))

    # create graph instance
    g <- graph.adjacency(score_mat[ind, ind], mode='undirected', weighted=TRUE)

    # get sum of edge weights for each vertex and rescale to range [-1, 1]
    edge_flux <- graph.strength(g)
    edge_flux <- edge_flux / max(abs(edge_flux))

    # color vertices based on net edge flux
    pal_neg <- colorRamp(c('purple', 'red'))
    pal_pos <- colorRamp(c('purple', 'blue'))

    V(g)$color <- sapply(edge_flux, FUN=function(x) {
        if (x < 0) {
            rgbcol <- pal_neg(abs(x))
        } else {
            rgbcol <- pal_pos(x)
        }
        rgb(rgbcol[1], rgbcol[2], rgbcol[3], 128, maxColorValue=255)
    })

    # color edges based on direction of change (red = neg, blue = pos)
    E(g)$color <- ifelse(E(g)$weight < 0, rgb(1,0,0,0.5), rgb(0,0,1,0.5))
    E(g)$weight <- abs(E(g)$weight)

    # remove low-weight edges (keep an average of 3 edges per node)
    num_edges <- 3 * length(V(g))

    if (num_edges < length(E(g))) {
        cutoff <- as.numeric(quantile(E(g)$weight, 1 - (num_edges / length(E(g)))))

        E(g)$weight[E(g)$weight < cutoff] <- 0
        g <- delete.edges(g, which(E(g)$weight==0))

        # remove nodes that become disconnected
        g <- delete.vertices(g, which(degree(g) == 0))
    }

    # relabel
    vlabels <- gene_info[match(names(V(g)), gene_info$gene_id),]$description

    # rescale edge weights
    E(g)$weight <- E(g)$weight - min(E(g)$weight)
    E(g)$weight <- ((E(g)$weight / max(E(g)$weight))) * 5

    layout_coords <- layout.fruchterman.reingold(g)

    V(g)$label.cex   <- 0.6
    V(g)$label.color <- '#333333'

    # network plot
    plot(g, layout=layout_coords, vertex.size=4, vertex.label=vlabels,
         vertex.label.dist=1,
         edge.width=E(g)$weight)

    # heatmap plot
    adj <- score_mat[ind,ind]

    heatmap.2(adj, trace='none', dendrogram='row', cexRow=30/nrow(adj),
              labRow=gene_info$description[ind], margins=c(6, 22),
              labCol=FALSE, col=pal,
              xlab='Gene', ylab='Gene', 
              key.title='', key.ylab='', keysize=2,
              key.xlab='Co-ex Conservation Score\n (HsTc / HsLm)')
}
```

```{r save_network, message=FALSE}
message("Saving difference network GraphML...")

# rescale edge weights to [-1, 1] and save network
#score_mat <- log1p(score_mat)
score_mat <- score_mat / max(abs(score_mat))

outfile <- file.path('..', output_datadir, paste0('coex_conservation_network_',
                                                  output_suffix, '.graphml'))

# annotations
annot <- cbind(gene_info, color=module_colors, cluster=module_labels)

g <- export_network_to_graphml(score_mat,
                               outfile,
                               threshold=0, max_edge_ratio=4,
                               nodeAttrDataFrame=annot)
```
